---
title: "Math 255 - Homework 5"
author: "Colin Pi"
date: "Due in class, Monday April 29"
output: pdf_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, prompt=TRUE, comment=NULL,tidy.opts=list(width.cutoff=60),tidy=TRUE,eval=TRUE, out.width = "60%")
library(ggplot2)
library(dplyr)
library(survey)
library(SDaA)
```

### Problem 1

$N_1 = 1000,\ N_2 = 1000,\ S_1 = 10,\ S_2 = 20,\ c_1 = 2,\ c_2 = 1$

**(a)**

c({$a_h$},n) = $c_0 + \sum\limits_{h=1}^{H}c_h(n\cdot a_h) = 0 + 2\cdot50 + 1\cdot50 = 150$\
V({$a_h$},n) = $\sum\limits_{h=1}^{H}(\dfrac{N_h}{N})^2(1-\dfrac{n\cdot a_h}{N_h})\dfrac{S^2_h}{n\cdot a_h} = (\dfrac{1000}{2000})^2(1-\dfrac{50}{1000})\dfrac{100}{50} + (\dfrac{1000}{2000})^2(1-\dfrac{50}{1000})\dfrac{1}{50}$ = 0.47975

**(b)**

$a_1 = \dfrac{N_1S_1/\sqrt{c_1}}{\sum\limits_{k=1}^H N_kS_k/\sqrt{c_k}} = \dfrac{1000\cdot10/\sqrt{2}}{1000\cdot10/\sqrt{2}+1000\cdot1/\sqrt{1}}$ = 0.8761007\

$a_2 = \dfrac{N_2S_2/\sqrt{c_2}}{\sum\limits_{k=1}^H N_kS_k/\sqrt{c_k}} = \dfrac{1000\cdot1/\sqrt{1}}{1000\cdot10/\sqrt{2}+1000\cdot1/\sqrt{1}}$ = 0.1238993\

**(c)**

150 = $\sum\limits_{h=1}^Hc_h(n\cdot a_h) = 2(n\cdot0.8761007) + 1(n\cdot0.1238993) = 1.876101n$\
n = $\dfrac{150}{1.876101} = 79.95305 = 79$

**(d)**

V({$a_h$},n) = $\sum\limits_{h=1}^{H}(\dfrac{N_h}{N})^2(1-\dfrac{n\cdot a_h}{N_h})\dfrac{S^2_h}{n\cdot a_h} = (\dfrac{1000}{2000})^2(1-\dfrac{69}{1000})\dfrac{100}{69} + (\dfrac{1000}{2000})^2(1-\dfrac{10}{1000})\dfrac{1}{10} = 0.3620688$

The budget is the same with (a), but the variance is lower than (a).

**(e)**

```{r}
c <- c(2,1)
a <- c(0.8761007, 0.1238993)
N <- c(1000,1000)
S <- c(10,1)

n_0 <- qnorm(0.975)^2*sum((N/sum(N))^2*(S^2/a))/1.36^2
n_0

n <- 64/(1+64/2000)
n

cost <- sum(c*round(63*a))
cost

var <- sum((N/sum(N))^2*(1-round(63*a)/N)*S^2/round(63*a))
var
```


$n_0 = \dfrac{1.96^2\cdot\sum\limits_{h=1}^H(\dfrac{N_h}{N})^2\dfrac{S^2_h}{a_h}}{1.36^2} = 63.45656 = 64$\
$n = \dfrac{n_0}{1+\dfrac{n_0}{2000}} = \dfrac{64}{1+\dfrac{64}{2000}} = 62.0155 = 63$\
c = $\sum\limits_{h=1}^{H}c_h(n\cdot a_h) = 118$\
$Var[\bar{y}_{str}] = \sum\limits_{h}(\dfrac{N_h}{N})^2(1-\dfrac{n\cdot a_h}{N_h})(\dfrac{S^2_h}{n\cdot a_h}) = 0.4605455$\

Both budget and variance is smaller than the scenario (a)

### Problem 2
Lohr textbook ch. 3 exercise 8.

$N_{phone} = 0.9N, N_{nonphone} = 0.1N, S_{phone} \approx S_{nonphone} = S$

**(a)**

$a_{phone} = \dfrac{\dfrac{0.9NS}{\sqrt{30}}}{\dfrac{0.9NS}{\sqrt{30}}+\dfrac{0.1NS}{\sqrt{40}}} = \dfrac{0.9\sqrt{40}}{0.9\sqrt{40}+0.1\sqrt{30}} = 0.9122215$\

$a_{nonphone} = \dfrac{\dfrac{0.1NS}{\sqrt{40}}}{\dfrac{0.9NS}{\sqrt{30}}+\dfrac{0.1NS}{\sqrt{40}}} = \dfrac{0.1\sqrt{30}}{0.9\sqrt{40}+0.1\sqrt{30}} = 0.08777855$\

$n = \dfrac{C-c_0}{\sum\limits_{h=1}^H c_ha_h} = \dfrac{20000-5000}{30\cdot0.9122215 + 40\cdot0.08777855} = 485.7861 \approx 485$\

$n_{phone} = n\cdot a_{phone} = 442.42 \approx 442$\

$n_{nonphone} = n\cdot a_{nonphone} = 42.5726 \approx 43$\

**(b)**

$a_{phone} = \dfrac{\dfrac{0.9NS}{\sqrt{10}}}{\dfrac{0.9NS}{\sqrt{10}}+\dfrac{0.1NS}{\sqrt{40}}} = \dfrac{0.9\sqrt{40}}{0.9\sqrt{40}+0.1\sqrt{10}} = 0.9473684$\

$a_{nonphone} = \dfrac{\dfrac{0.1NS}{\sqrt{40}}}{\dfrac{0.9NS}{\sqrt{10}}+\dfrac{0.1NS}{\sqrt{40}}} = \dfrac{0.1\sqrt{10}}{0.9\sqrt{40}+0.1\sqrt{10}} = 0.05263158$\

$n = \dfrac{C-c_0}{\sum\limits_{h=1}^H c_ha_h} = \dfrac{20000-5000}{10\cdot0.9473684 + 40\cdot0.05263158} = 1295.455 \approx 1295$\

$n_{phone} = n\cdot a_{phone} = 1226.842 \approx 1227$\

$n_{nonphone} = n\cdot a_{nonphone} = 68.1579 \approx 68$

### Problem 3
Lohr textbook ch. 3 exercise 22(a-b)\

$c_1 = c_2, \dfrac{N_1}{N} = 0.4, n = 2000$

**(a)**

Since $c_1 = c_2$, it is Neyman allocation.\

$a_1 = \dfrac{0.4N\sqrt{0.9\cdot0.1}}{0.4N\sqrt{0.9\cdot0.1} + 0.6N\sqrt{0.97\cdot0.03}} = 0.539684$\

$a_2 = \dfrac{0.6N\sqrt{0.97\cdot0.03}}{0.4N\sqrt{0.9\cdot0.1} + 0.6N\sqrt{0.97\cdot0.03}} = 0.460316$\

$n_1 = n\cdot a_1 = 1079.368 \approx 1079$\

$n_2 = n\cdot a_2 = 920.632 \approx 921$\

**(b)**

In that N is a huge number, let's assume $1-\dfrac{n_h}{N_h} \approx 1$.

- Proportional Allocation

$V[\hat{p}_{str}] = \sum\limits_{h=1}^H(\dfrac{N_h}{N})^2\dfrac{p_h(1-p_h)}{n_h-1} = 0.4^2(\dfrac{0.9\cdot0.1}{0.4\cdot2000-1}) + 0.6^2(\dfrac{0.97\cdot0.03}{0.6\cdot2000-1}) = 2.675981 \cdot 10^{-5}$

- Optimal Allocation

$V[\hat{p}_{str}] = \sum\limits_{h=1}^H(\dfrac{N_h}{N})^2\dfrac{p_h(1-p_h)}{n_h-1} = 0.4^2(\dfrac{0.9\cdot0.1}{1079-1}) + 0.6^2(\dfrac{0.97\cdot0.03}{921-1}) = 2.675981 \cdot 10^{-5} = 2.474503\cdot10^{-5}$

- SRS

$\hat{p} = \sum\limits_{h} \dfrac{N_h}{N}p_h = 0.4\cdot0.10 + 0.6\cdot0.03 = 0.058$\

$V[\hat{p}_{ssr}] = (1-\dfrac{n}{N})\dfrac{\hat{p}(1-\hat{p})}{n-1} \approx \dfrac{\hat{p}(1-\hat{p})}{n-1} = 2.733167\cdot10^{-5}$

### Problem 4
Lohr textbook ch. 4 exercise 1. 

**(a)**

$x_i$ = television news broadcasts time in day $i$\
$y_i$ = time in television news broadcasts devoted to sports in day $i$\
$\hat{p}_r = \hat{B} = \dfrac{t_y}{t_x}$

**(b)**

$x_i$ = number of fish an angler $i$ caught in August\
$y_i$ = number of fish caught by angler $i$ in a lake in August\
$\hat{\bar{y}}_r = \hat{B}\bar{x}_U = \bar{y}\dfrac{\bar{x}_U}{\bar{x}}$, where B = $\dfrac{\bar{y}}{\bar{x}} = \dfrac{t_y}{t_x}$

**(c)**

$x_i$ = spending of undergraduate student $i$ in fall term\
$y_i$ = spending of undergraduate student $i$ in fall term for textbook\
$\hat{\bar{y}}_r = \hat{B}\bar{x}_U = \bar{y}\dfrac{\bar{x}_U}{\bar{x}}$, where B = $\dfrac{\bar{y}}{\bar{x}} = \dfrac{t_y}{t_x}$

**(d)**

$x_i$ = weight of chicken $i$\
$y_i$ = weight of usable meat of chicken $i$\
$t_{yr} = \hat{B}t_x = \dfrac{\bar{y}}{\bar{x}}t_x = \bar{y}\dfrac{t_x}{\bar{x}}$

### Problem 5
Lohr textbook ch. 4 exercise 2

```{r}
y <- c(10,7,13,17,8,1,15,7,4)
x <- c(13,7,11,12,4,3,11,3,5)
n <- 3
N <- 9
```

**(a)**

```{r}
t.x <- sum(x)
t.y <- sum(y)
s.x <- sd(x)
s.y <- sd(y)
r <- cor(x,y)
b <- t.y/t.x
```

$t_x$ = `r t.x`\
$t_y$ = `r t.y`\
$S_x$ = `r s.x`\
$S_y$ = `r s.y`\
$R$ = `r r`\
$B$ = `r b`

**(b)**

```{r}
reps <- 10000
results <- data.frame(run = 1:reps, t.srs = NA, t.ratio = NA)
set.seed(124)

for (i in 1:reps){
  s <- sample(1:N, n, replace=F)
  y.samp <- y[s]
  x.samp <- x[s]
  results$t.srs[i] <- N*mean(y.samp)
  results$t.ratio[i] <- sum(y.samp)/sum(x.samp)*t.x
}
```


**(c)**

```{r, fig.align='center'}
par(mfrow=c(1,2))
hist(results$t.srs, main = "SRS", xlab = "")
hist(results$t.ratio, main = "Ratio", xlab = "")
par(mfrow=c(1,1))
```

**(d)**

```{r}
knitr::kable(data.frame(Statistics = c("Mean","Variance","Bias"), 
                        yr = c(mean(results$t.ratio, na.rm = T), var(results$t.ratio, na.rm = T), mean(results$t.ratio, na.rm = T) - t.y), 
                        ysrs = c(mean(results$t.srs, na.rm = T), var(results$t.srs, na.rm = T), mean(results$t.srs, na.rm = T) - t.y)))
```

$\hat{t}_{yr}$ has expected value close to $t_x$, so the bias is close to 0 (`r mean(results$t.ratio, na.rm = T) - t.y`) . But the expected value of $\hat{t}_{y,\ SRS}$ is closer than that that of $\hat{t}_{yr}$ with bias of `r mean(results$t.srs, na.rm = T) - t.y`. The variance of $\hat{t}_{yr}$ is almost 1/2 of $\hat{t}_{y,\ SRS}$. 

**(e)**

$\bar{x}_U$ = `r t.x/9`

```{r}
(1-3/10)*(1/(3*7.666667))*(b*s.x^2-r*s.x*s.y)
```


$$
Bias(\hat{\bar{y}}_r) = (1-\dfrac{n}{N})\dfrac{1}{n\bar{x}_U}(BS^2_x-RS_xS_y)
$$
$Bias(\hat{\bar{y}}_r) = (1-3/10)\dfrac{1}{3\cdot 6.9}(1.188406\cdot4.092676^2 - 0.8152062\cdot4.092676\cdot5.182771) = 0.08840055$\

$Bias(\hat{t}_{yr}) \approx NBias(\hat{\bar{y}}_r) = 0.8840055$\

It is close to the bias calculated in (c) (`r mean(results$t.ratio, na.rm = T) - t.y`).

### Problem 6
Lohr textbook ch. 4 exercise 8(a,b,d). For part (d), ignore the regression estimator when answering the question.

**(a)**

```{r, fig.align='center'}
ggplot(data = agsrs, aes(x = farms87, y = acres92)) +
  geom_point() +
  ggtitle("acres92 vs. farms87")
```

**(b)**

$y_i$ = number of acres devoted to farming in county $i$ in 1992\
$x_i$ = number of farms in county $i$ in 1987\
$\hat{t}_{yr} = Bt_x = \dfrac{\bar{y}}{\bar{x}}t_x = \dfrac{t_x}{\bar{x}}\bar{y}$ = 960,155,061

```{r}
agsrs$n <- nrow(agsrs)
agsrs$N <- 3078
agsrs$wts <- agsrs$N/agsrs$n
design6.srs <- svydesign(id=~1, fpc=~N, weights=~wts, data = agsrs)

ratio.farms <- svyratio(~acres92,~farms87, design6.srs)
tx.farms <- 2087759
ty.farms <- predict(ratio.farms, tx.farms)
ty.farms
confint(ratio.farms, df=degf(design6.srs))*tx.farms
```

**(d)**

```{r}
ratio.acres <- svyratio(~acres92,~acres87, design6.srs)
tx.acres <- 964470625
ty.acres <- predict(ratio.acres, tx.acres)
ty.acres
confint(ratio.acres, df=degf(design6.srs))*tx.acres

cor(agsrs$acres92, agsrs$acres87)
cor(agsrs$acres92, agsrs$farms87)
```

Ratio estimation with auxiliary variable `acres87` gives the most precision (SE = 5546162). SE[$\hat{B}$] decreases as $\hat{R}$ increases, or in other words, we can get more precise ratio estimation as the auxiliary variable gets more correlated with our variable of interest. As noted above, `acres92` exhibits a strong correlation with `acres87` ($\hat{R}$ = 0.995806) while not with `farms87` ($\hat{R}$ = 0.05965), and this difference accounts why using `acres87` gives more precise result than using `farms87`.

### Problem 7
Lohr textbook ch. 4 exercise 10(a,b). Data is found:

```{r}
cherry <- read.csv("http://math.carleton.edu/kstclair/data/cherry.csv")
```

**(a)**

```{r, fig.align='center'}
ggplot(data = cherry, aes(x = diameter, y = volume)) +
  geom_point() +
  ggtitle("Volume vs. Diameter")
```

**(b)**

$y_i$ = volume of black cherry tree $i$\
$x_i$ = diameter of black cherry tree $i$\
$\hat{t}_{yr} = Bt_x = \dfrac{\bar{y}}{\bar{x}}t_x = \dfrac{t_x}{\bar{x}}\bar{y} = \dfrac{41,835}{13.24839}\cdot30.17097$ = 95272.16\
$SE[\hat{t}_{yr}] = N\sqrt{(1-\dfrac{n}{N})(\dfrac{\bar{x}_U}{\bar{x}})^2\dfrac{s^2_e}{n}} = 2967\sqrt{(1-\dfrac{31}{2967})(\dfrac{14.1001}{13.24839})^2\dfrac{s^2_e}{31}}$\
= $2967\sqrt{(1-\dfrac{31}{2967})(\dfrac{14.1001}{13.24839})^2\dfrac{94.05287}{31}}$ = 5471.434\
$s^2_e = s^2_y + \hat{B}^2s^2_x - 2\hat{B}\hat{R}s_ys_x = 270.2028 + 2.277331^2\cdot9.847914-2\cdot2.277331\cdot0.9671194\cdot16.43785\cdot3.138139$ = 94.05287

95% CI: $\hat{t}_{yr} \pm qt_{0.975,\ df=30}\cdot SE[\hat{t}_{yr}]$ = (84098.0, 106446.3)

```{r}
t.x.7 <- 41835
b.7 <- sum(cherry$volume)/sum(cherry$diameter)
t.y.hat.7 <- b.7*t.x.7
t.y.hat.7

xbar_u <- t.x.7/2967

s.e.7_2 <- var(cherry$volume) + b.7^2*var(cherry$diameter) - 2*b.7*cor(cherry$volume,cherry$diameter)*sd(cherry$volume)*sd(cherry$diameter)

se.7 <- 2967*sqrt((1-31/2967)*(xbar_u/mean(cherry$diameter))^2*s.e.7_2/31)

t.y.hat.7 - qt(c(0.975,0.025), df = 30)*se.7

cherry$n <- nrow(cherry)
cherry$N <- 2967
cherry$wts <- cherry$N/cherry$n
design7.srs <- svydesign(id=~1, fpc=~N, weights=~wts, data = cherry)

ratio.cherry <- svyratio(~volume,~diameter, design7.srs)
tx.cherry <- t.x.7
ty.cherry <- predict(ratio.cherry, tx.cherry)
ty.cherry
confint(ratio.cherry, df=degf(design7.srs))*tx.cherry
```

### Problem 8

Assuming $\dfrac{\bar{x}_U}{\bar{x}}$,\

$SE[\bar{y}_r] = \sqrt{(1-\dfrac{n}{N})(\dfrac{S^2_y+B^2S^2_x-2BRS_xS_y}{n})} < \sqrt{(1-\dfrac{n}{N})\dfrac{S^2_y}{n}}$\

If we cancel out some terms,\

$BS_x = \dfrac{\bar{y}}{\bar{x}}S_x < 2RS_y$\

We know that CV(x) = $\dfrac{S_x}{\bar{x}}$, CV(y) = $\dfrac{S_y}{\bar{y}}$. If we rearrange the equation above,\

$\dfrac{S_x}{\bar{x}} < 2R\dfrac{S_y}{\bar{y}} \rightarrow \dfrac{CV(x)}{2CV(y)} < R$

