---
title: "Math 255 - Homework 4"
author: "Colin Pi"
date: "Due in class, Friday April 19"
output: pdf_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, prompt=TRUE, comment=NULL,tidy.opts=list(width.cutoff=60),tidy=TRUE,eval=TRUE, out.width = "60%")
library(ggplot2)
library(dplyr)
library(survey)
library(SDaA)
```

### Problem 1
Lohr textbook ch. 3 exercise 5.

**(a)** Selected scholars in American Council of Learned Societies in seven disciplines who answered to the survey.

**(b)**

```{r}
Nh <- c(9100, 1950, 5500, 10850, 2100, 5500, 9000)
N <- sum(Nh)
p.hats <- c(0.37, 0.23, 0.23, 0.29, 0.19, 0.43, 0.41)
phat.str.1 <- sum(Nh/N*p.hats)
phat.str.1

nh <- c(636, 451, 481, 611, 493, 575, 588)
var.str <- sum((Nh/N)^2*(1-nh/Nh)*p.hats*(1-p.hats)/(nh-1))
se.str.1 <- sqrt(var.str)
se.str.1
```

$\hat{p}_{str} = \sum\limits_{h=1}^{7} \dfrac{N_h}{N}\hat{p}_h = \dfrac{9100}{44000}\cdot0.37+...+\dfrac{9000}{44000}\cdot0.41$ = `r phat.str.1`\

$SE[\hat{p}_{str}] = \sqrt{\sum\limits_{h=1}^{7}(\dfrac{N_h}{N})^2(1-\dfrac{n_h}{N_h})\dfrac{\hat{p}(1-\hat{p})}{n_h-1}}$\
= $\sqrt{(\dfrac{9100}{44000})^2(1-\dfrac{636}{9100})\dfrac{0.37(1-0.37)}{635} +...+ (\dfrac{9000}{44000})^2(1-\dfrac{588}{9000})\dfrac{0.41(1-0.41)}{587}}$ = `r se.str.1`

### Problem 2
Lohr textbook ch. 3 exercise 9.

```{r}
agstrat$N <- recode(agstrat$region, NC = 1054, NE = 220, S = 1382, W = 422)
agstrat %>% group_by(region) %>% summarize(min(N), max(N))

agstrat <- agstrat %>% group_by(region) %>% mutate(n = n())
agstrat %>% group_by(region) %>% summarize(min(n),max(n)) # check

agstrat$wts <- agstrat$N/agstrat$n
agstrat %>% group_by(region) %>% summarize(min(wts),max(wts)) #check

design.strat <- svydesign(id=~1, fpc=~N, weights=~wts, strata=~region, data=agstrat)
```

**(a)**

```{r, fig.align='center'}
boxplot(acres87~region, data = agstrat, horizontal = TRUE, main = "Number of acres devoted to farms, 1987", xlab = "Acres")
svymean(~acres87, design.strat)
confint(svymean(~acres87, design.strat), df = degf(design.strat))
```

$\bar{y}_{ssr}\ (301954)$ > $\bar{y}_{str}\ (298547)$\
$SE[\bar{y}_{ssr}]\ (18914)$ > $SE[\bar{y}_{str}]\ (16293)$\
$CI_{ssr} = (264733, 339174.5)$ is wider than $CI_{str} = (266482.4, 330611.8)$ 

**(b)**

```{r, fig.align='center'}
boxplot(farms92~region, data = agstrat, horizontal = TRUE, main = "Number of farms, 1992", xlab = "Farms")
svymean(~farms92, design.strat)
confint(svymean(~farms92, design.strat),df = degf(design.strat))
```

$\bar{y}_{ssr}\ (599.06)$ < $\bar{y}_{str}\ (637.16)$\
$SE[\bar{y}_{ssr}]\ (22.062)$ < $SE[\bar{y}_{str}]\ (24.277)$\
$CI_{ssr} = (555.6426, 642.4774)$ is narrower than $CI_{str} = (589.3853, 684.9422)$. 

**(c)**

```{r, fig.align='center'}
boxplot(largef92~region, data = agstrat, horizontal = TRUE, main = "Number of farms with 1000 acres or more, 1992", xlab = "Farms")
svymean(~largef92, design.strat)
confint(svymean(~largef92, design.strat),df = degf(design.strat))
```

$\bar{y}_{ssr} = 56.593$ < $\bar{y}_{str} = 56.698$ (approximately the same)\
$SE[\bar{y}_{ssr}] = 3.9904$ > $SE[\bar{y}_{str}] = 3.5577$\
$CI_{ssr} = (48.77239, 64.41428)$ is wider than $CI_{str} = (49.69636, 63.69954)$. 

**(d)**

```{r, fig.align='center'}
boxplot(smallf92~region, data = agstrat, horizontal = TRUE, main = "Number of farms with 9 acres or fewer, 1992", xlab = "Farms")
svymean(~largef92, design.strat)
confint(svymean(~largef92, design.strat),df = degf(design.strat))
svymean(~smallf92, design.strat)
confint(svymean(~smallf92, design.strat),df = degf(design.strat))
```

$\bar{y}_{ssr} = 46.823$ < $\bar{y}_{str} = 56.863$\
$SE[\bar{y}_{ssr}] = 3.6375$ < $SE[\bar{y}_{str}] = 7.2014$\
$CI_{ssr} = (39.69387, 53.95279)$ is narrower than $CI_{str} = (42.69033, 71.03526)$. 

### Problem 3
Lohr textbook ch. 3 exercise 15.

**(a)**

Advantage: proportional allocation provides the most precise result when within variance of all strata are similar.\
Disadvantage: If the within variance varies on strata, proportional allocation is not the best way to produce the most precise result (optimal allocation is better than proportional allocation in this case). 

**(b)**

```{r}
Nh <- c(190, 407, 811)
ybarh <- c(3.925, 3.938, 3.942)
N <- sum(Nh)
ybar.str.3 <- sum(Nh/N*ybarh)

sh <- c(0.037, 0.052, 0.070)
nh <- c(21,14,22)
se.str.3 <- sqrt(sum((Nh/N)^2*(1-nh/Nh)*sh^2/nh))
se.str.3
ci.3 <- ybar.str.3 - se.str.3*qt(c(0.975,0.025),21+14+22-3)
ci.3
```

$\bar{y}_U = \sum\limits_{h = low}^{upper} \dfrac{N_h}{N}\bar{y}_h = \dfrac{190}{1408}\cdot 3.925+\dfrac{407}{1408}\cdot 3.938+\dfrac{811}{1408}\cdot 3.942$ = `r ybar.str.3`\

$SE[\bar{y}_U] = \sqrt{\sum\limits_{h=low}^{upper}(\dfrac{N_h}{N})^2(1-\dfrac{n_h}{N_h})\dfrac{s_h^2}{n_h}}$\

= $\sqrt{(\dfrac{190}{1408})^2\cdot(1-\dfrac{21}{190})\dfrac{0.037^2}{21}+(\dfrac{407}{1408})^2\cdot(1-\dfrac{14}{407})\dfrac{0.053^2}{14}+(\dfrac{811}{1408})^2\cdot(1-\dfrac{22}{811})\dfrac{0.070^2}{22}}$ = `r se.str.3`\

95% CI: $\bar{y}_U \pm qt_{0.975, df = 57-3}\cdot SE[\bar{y}_U]$ = (`r ci.3[1]`, `r ci.3[2]`)

**(c)**

We can answer this answer by estimating the difference in means of log prices among the three strata.

- Low v. Middle

```{r}
diff.1 <- 3.925-3.938

se.1 <- sqrt(sum((1-nh[1:2]/Nh[1:2])*sh[1:2]^2/nh[1:2]))
se.1

ci.3.1 <- diff.1 - se.1*qnorm(c(0.975,0.025)) # what df should I use?
```

$$
\bar{y}_{low} - \bar{y}_{middle} \pm qt_{0.975,\ df=}\cdot\sqrt{Var[\bar{y}_{low}] + Var[\bar{y}_{middle}]} = -0.013 \pm 1.96\cdot\sqrt{0.000057985 + 0.000185}
$$

95% CI: (`r ci.3.1[1]`, `r ci.3.1[2]`)\
We are 95% confident that the log price of low income stratum is `r ci.3.1[1]` to `r ci.3.1[2]` higher than middle income stratum in average. We can't conclude that the log price of low income stratum is different from that of middle income stratum.

- Middle v. Upper

```{r}
diff.2 <- 3.938-3.942

se.2 <- sqrt(sum((1-nh[2:3]/Nh[2:3])*sh[2:3]^2/nh[2:3]))
se.2

ci.3.2 <- diff.2 - se.2*qnorm(c(0.975,0.025)) # what df should I use?
ci.3.2
```

$$
\bar{y}_{middle} - \bar{y}_{upper} \pm qt_{0.975,\ df=}\cdot\sqrt{Var[\bar{y}_{middle}] + Var[\bar{y}_{upper}]} = -0.004 \pm 1.96\cdot\sqrt{0.0001864991 + 0.0002166853}
$$

95% CI: (`r ci.3.2[1]`, `r ci.3.2[2]`)\
We are 95% confident that the log price of middle income stratum is `r ci.3.2[1]` to `r ci.3.2[2]` higher than upper income stratum in average. We can't conclude that the log price of middle income stratum is different from that of upper income stratum.

- High v. low

```{r}
diff.3 <- 3.925-3.942

se.3 <- sqrt(sum((1-nh[c(1,3)]/Nh[c(1,3)])*sh[c(1,3)]^2/nh[c(1,3)]))
se.3

ci.3.3 <- diff.3 - se.3*qnorm(c(0.975,0.025)) # what df should I use?
ci.3.3
```

$$
\bar{y}_{low} - \bar{y}_{upper} \pm qt_{0.975,\ df=}\cdot\sqrt{Var[\bar{y}_{low}] + Var[\bar{y}_{upper}]} = -0.017 \pm 1.96\cdot\sqrt{0.000057985 + 0.0002166853}
$$

95% CI: (`r ci.3.3[1]`, `r ci.3.3[2]`)\
We are 95% confident that the log price of low income stratum is `r ci.3.3[1]` to `r ci.3.3[2]` higher than upper income stratum in average. We can't conclude that the log price of low income stratum is different from that of upper income stratum.

### Problem 4
Lohr textbook ch. 3 exercise 16. (Data in SDaA.)

```{r}
otters$N <- recode(otters$habitat, `1` = 89, `2` = 61, `3` = 40, `4` = 47)
otters %>% group_by(habitat) %>% summarize(min(N), max(N))

otters <- otters %>% group_by(habitat) %>% mutate(n = n())
otters %>% group_by(habitat) %>% summarize(min(n),max(n)) # check

otters$wts <- otters$N/otters$n
otters %>% group_by(habitat) %>% summarize(min(wts),max(wts)) #check

design4.strat <- svydesign(id=~1, fpc=~N, weights=~wts, strata=~habitat, data=otters)
```

**(a)**

```{r}
svytotal(~holts, design4.strat)
```

$\hat{t}_{str}$ = 984.71\
$SE[\hat{t}_{str}]$ = 73.921

**(b)**

The study area is divided into stratum based on the predominant terrain type. In other words, some of the sections may exhibit characteristics of more than one classification (for example we can see Cliffs and Agriculture in certain section). So classifying such sections heaviliy relies on the researchers' judgement, possibly resulting in selection bias. Also, some of the dens can be either abandoned or belonged to other animals, implying that the study is not also free from measurement error. 

### Problem 5

**(a)** Households that are not listed in the county's telephone number have no chance to be sampled, so it is not free from selection bias. Also, the survey is not free from nonresponse issue as well. 

**(b)**

```{r}
radon <- read.csv("http://math.carleton.edu/kstclair/data/radon.csv")
options(survey.lonely.psu="remove")

radon$wts <- radon$popsize/radon$sampsize
head(radon %>% group_by(countyname) %>% summarize(min(wts),max(wts))) #check

design5.strat <- svydesign(id=~1, fpc= ~popsize, weights = ~wts, strata=~countyname, data = radon)
svymean(~radon, design5.strat)
```

$\hat{\bar{y}}_{str}$ = 4.8986 pCi/L\
$SE[\hat{\bar{y}}_{str}]$ = 0.1543 pCi/L

**(c)**

```{r}
htf4radon <- ifelse(radon$radon >= 4, 1, 0)
update(design5.strat, htf4radon = htf4radon)
svytotal(~htf4radon, design5.strat)
confint(svytotal(~htf4radon, design5.strat), df = degf(design5.strat))
```

$\hat{t}_{radon>4 pCi/L}$ = 722781 households\
95% CI: (667632.1, 777930.4) households

**(d)**

```{r}
head(radon %>% group_by(countyname) %>% arrange(sampsize))
```

Mahomen, Murray, and Wilkin county only have 1 home sampled.

**(e)**

```{r, fig.align='center'}
svyby.out <- svyby(~radon, ~countyname, design5.strat, svymean)
ggplot(svyby.out, aes(y=countyname, x=radon)) +
  geom_point() +
  geom_errorbarh(aes(xmax=radon + se, xmin = radon - se))
```

**(f)**

```{r}
head(svyby.out %>% arrange(se))
head(svyby.out %>% arrange(desc(se)))
```

The county with the largest SE for estimating the mean randon level of homes is Marshall county. The county with the smallest SE for estimating the mean randon level of homes is Sherburne county. 

```{r}
radon %>% 
  filter(countyname %in% c("Marshall","Sherburne")) %>% 
  group_by(countyname) %>% 
  summarize(n = n(), sd = sd(radon))

16.2/0.557

5.389193/0.1855169
```

$$SE[\hat{\bar{y}}] = \sqrt{(1-\dfrac{n_h}{N_h})}\dfrac{s_h}{\sqrt{n_h}}$$
SE is proportional to sample standard deviation and inversely proportional to square root of sample size. 9 homes are sampled from both counties, but the sample standard deviation of Marshall county is almost 30 times that of Sherburne county, which is close to the ratio of standard errors between the two counties (let's ignore fpc as population size is big). In sum, significant difference in sample standard deviation between the two counties (while the number of the sample collected are the same) is biggest reason why the SE's of these counties are either biggest or smallest.

### Problem 6
Revisit problem 2 above. Compute and interpret the design effect using the survey package for each of four estimates computed for exercise 9. Which estimate has the smallest DEff and which has the largest? Use the EDA (graphs) you produced for problem 2 to explain why these variables have the smallest and largest DEff.

```{r}
svymean(~acres87, design.strat, deff=T)
svymean(~farms92, design.strat, deff=T)
svymean(~largef92, design.strat, deff=T)
svymean(~smallf92, design.strat, deff=T)
```

$$
Var(\hat{t}_{str}) = (1-\dfrac{n}{N})\dfrac{N}{n}(SSW + \sum\limits_{h}s_h^2)
$$

$$
Var(\hat{t}_{ssr}) = (1-\dfrac{n}{N})N^2\dfrac{SSB+SSW}{n(N-1)}
$$
$$
DEff = \dfrac{Var(\hat{t}_{str})}{Var(\hat{t}_{ssr})} = \dfrac{(1-\dfrac{n}{N})\dfrac{N}{n}(SSW + \sum\limits_{h}s_h^2)}{(1-\dfrac{n}{N})N^2\dfrac{SSB+SSW}{n(N-1)}} = \dfrac{SSW + \sum\limits_{h}s_h^2}{\dfrac{N(SSB+SSW)}{N-1}} = \dfrac{SSW + \sum\limits_{h}s_h^2}{NS^2}
$$

The equation above suggests that DEff gets bigger as the proportion of SSW + $\sum\limits_{h}s_h^2$ over SST (approximately, as SST = $(N-1)S^2$) gets bigger. In other words, DEff gets larger if the relative portion of SSB in SST gets smaller (SSW + SSB = SST). 
\
`acres87` has the smallest DEff, and `smallf92` has the largest DEff. EDB illustrates that the relatitve portion of SSB on SST in the case of `acres87` is big because we can observe huge difference in values among different strata; however, relative portion of SSB on SST is small in `smallf92` as we only see small difference in values among different strata. 

```{r}
acres.aov <- aov(acres87~region, data=agstrat)
anova(acres.aov)

smallf.aov <- aov(smallf92~region, data=agstrat)
anova(smallf.aov)
```

The ANOVA output also supports the claim that the relative size of SSB to SST in `acres87` case is significantly bigger than `smallf92` case. Approximately, SSB/SST in `acres87` case is $\dfrac{6.9528e+12}{(6.9528e+12) + (2.5937e+13)}$ = 0.2113968 but only $\dfrac{217384}{217384+5035679}$ = 0.04138233 in `smallf92` case. 

### Problem 7
Lohr textbook ch. 3 exercise 35 parts (a)-(d).

**(a)**

```{r}
pop <- read.csv("http://math.carleton.edu/kstclair/data/baseball.csv", header=FALSE, na.strings = c("NA"," ", "."))
names(pop) <- c("team","league","player","salary","POS","G","GS",
"InnOuts","PO","A","E","DP","PB","GB","AB","R","H",
"SecB","ThiB","HR","RBI","SB","CS","BB","SO","IBB",
"HPB","SH","SF","GIDP")

table(pop$team) # all roughly the same size populations
pop$logsal <- log(pop$salary)

pop$N <- recode(pop$team, ANA = 26, ARI=28, ATL=28, BAL=25, BOS=27, CHA=26, CHN=29, CIN=27, CLE=28, COL=27, DET=26, FLO=26, HOU=25, KCA=27, LAN=24, MIL=25, MIN=25, MON=28, NYA=29, NYN=26, OAK=27, PHI=25, PIT=27, SDN=26, SEA=27, SFN=28, SLN=26, TBA=26, TEX=27, TOR=26)

head(pop %>% group_by(team) %>% summarize(min(N), max(N)))

set.seed(30) # put your favorite large integer here
baseball.strat <- pop %>% group_by(team) %>% sample_n(size=5) %>% ungroup()
str(baseball.strat)
```

I used 5 samples from each strata using SRS (number of sample is constant (5) because population size is roughly same across the strata). 

**(b)**

```{r}
baseball.strat$wts <- baseball.strat$N/5

design7.strat <- svydesign(id=~1, fpc=~N, weights=~wts, strata=~team, data = baseball.strat)

svymean(~logsal, design7.strat)
confint(svymean(~logsal, design7.strat), df=degf(design7.strat))
```

$\bar{\hat{logsal}}$ = 13.883\
95% CI: (13.70935, 14.05757)

**(c)**

```{r}
pitcher <- ifelse(baseball.strat$POS == "P", 1,0)
update(design7.strat, pitcher = pitcher)

svymean(~pitcher, design7.strat)
confint(svymean(~pitcher, design7.strat), df=degf(design7.strat))
```

$\hat{p}_{pitcher}$ = 0.44918\
95% CI: (0.3771607, 0.5212082)

**(d)**

```{r}
knitr::kable(data.frame(logsal = c("Mean", "SE", "CI"), ssr = c(13.982, 0.095, "(13.79421 14.16963)"), str = c(13.883, 0.0879, "(13.70935, 14.05757)")))

knitr::kable(data.frame(pitcher = c("Mean", "SE", "CI"), ssr = c(0.493333, 0.0369, "(0.420412591, 0.56625408)"), str = c(0.44918, 0.0364, "(0.3771607, 0.5212082)")))
```

The estimates from stratified sampling is smaller than those from SSR. Also the SE's are smaller than SSR. So, the CI's from stratified sampling is narrower than those from SSR. 
